---
output: html_document
editor_options: 
  chunk_output_type: inline
---
```{r}
library(tidyverse)
library(epifitter)
```

```{r}
browseURL("https://alvesks.github.io/epifitter/articles/fitting.html")
browseURL("https://emdelponte.github.io/epidemiology-R/temporal-analysis.html#model-fitting")
```

```{r}
dpcL <- sim_logistic(
  N = 100, # duration of the epidemics in days
  y0 = 0.01, # disease intensity at time zero
  dt = 10, # interval between assessments
  r = 0.1, # apparent infection rate
  alpha = 0.2, # level of noise
  n = 7 # number of replicates
)
```

```{r}
dpcL
```

```{r}
dpcL %>% 
  ggplot()+
  aes(time, y, group = replicates) +
  geom_point(aes(time, random_y), shape = 1) + # plot the replicate values
  geom_point(color = "steelblue", size = 2) +
  geom_line(color = "steelblue") +
  labs(
    title = "Simulated 'complete' epidemics of sigmoid shape",
    subtitle = "Produced using sim_logistic()"
  )+
  theme_minimal()
```

```{r}
f_lin <- fit_lin(time = dpcL$time,  y = dpcL$random_y)
```

```{r}
f_lin
```

```{r}
head(f_lin$data)
```

Plot of predictions

The plot_fit() produces, by default, a panel of plots depicting the observed and predicted values by all fitted models. The arguments pont_size and line_size that control for the size of the dots for the observation and the size of the fitted line, respectively.

```{r}
plot_lin <- plot_fit(f_lin,
  point_size = 2,
  line_size = 1
) 

# Default plots
plot_lin 
```

Two-parameters

The fit_nlin() function uses the Levenberg-Marquardt algorithm for least-squares estimation of nonlinear parameters. In addition to time and disease intensity, starting values for y0 and r should be given in the starting_par argument. The output format and interpretation is analogous to the fit_lin().

NOTE: If you encounter error messages saying “matrix at initial parameter estimates”, try to modify the starting values for the parameters to solve the problem.

```{r}
f_nlin <- fit_nlin(
  time = dpcL$time,
  y = dpcL$random_y,
  starting_par = list(y0 = 0.01, r = 0.03)
)

```

```{r}
f_nlin
```

```{r}
plot_fit(f_nlin) +
  theme_minimal()#changing plot theme
```

# Estimating K (maximum disease)

In many epidemics the last measure (final time) of a DPC does not reach the maximum intensity and, for this reason, estimation of maximum asymptote (carrying capacity K) may be necessary. The fin_lin2() provides an estimation of K in addition to the estimates provided by fit_lin().

Before demonstrating the function, we can transform our simulated data by creating another variable with y_random2 with maximum about 0.8 (80%). Simplest way is to multiply the y_random by 0.8.

```{r}
dpcL %>% 
  mutate(random_y = random_y * 0.8) -> dpcL2
dpcL2
```

Then we run the fit_nlin2() for the new dataset.

```{r}
f_nlin2 <- fit_nlin2(
  time = dpcL2$time,
  y = dpcL2$random_y,
  starting_par = list(y0 = 0.01, r = 0.2, K =  0.6)
)
f_nlin2

```

```{r}
plot_fit(f_nlin2)
```

# Fit models to multiple DPCs

Most commonly, there are more than one epidemics to analyse either from observational or experimental studies. When the goal is to fit a common model to all curves, the fit_multi() function is in hand. Each DPC needs an unique identified to further combined in a single data frame.

Data
Let’s use the sim_ family of functions to create three epidemics and store the data in a single data.frame. The Gompertz model was used to simulate these data. Note that we allowed to the y0 and r parameter to differ the DPCs. We should combine the three DPCs using the bind_rows() function and name the identifier (.id), automatically created as a character vector, for each epidemics as ‘DPC’.

```{r}
epi1 <- sim_gompertz(N = 60, y0 = 0.001, dt = 5, r = 0.1, alpha = 0.4, n = 4)
epi2 <- sim_gompertz(N = 60, y0 = 0.001, dt = 5, r = 0.12, alpha = 0.4, n = 4)
epi3 <- sim_gompertz(N = 60, y0 = 0.003, dt = 5, r = 0.14, alpha = 0.4, n = 4)

multi_epidemic <- bind_rows(epi1,
  epi2,
  epi3,
  .id = "DPC"
)
head(multi_epidemic)
```

We can visualize the three DPCs in a same plot

```{r}
multi_epidemic %>% 
  ggplot()+
  aes(time, y, col = DPC)+
  geom_point(size =2)+
  geom_line()+
  theme_minimal() +
  labs(x = "Time", y = "Proportion disease") -> p_multi
```

```{r}
p_multi
```

```{r}
p_multi +
  facet_wrap(~ DPC, ncol = 1)
```

Using fit_multi() 

fit_multi() requires at least four arguments: time, disease intensity (as proportion), data and the curve identifier (strata_cols). The latter argument accepts one or more strata include as c("strata1",strata2"). In the example below, the stratum name is DPC, the name of the variable.

By default, the linear regression is fitted to data but adding another argument nlin = T, the non linear regressions is fitted instead.

```{r}
multi_fit <- fit_multi(
  time_col = "time",
  intensity_col = "random_y",
  data = multi_epidemic,
  strata_cols = "DPC"
)
```

```{r}
head(multi_fit$Parameters, n=Inf)


```

```{r}
multi_fit2 <- fit_multi(
  time_col = "time",
  intensity_col = "random_y",
  data = multi_epidemic,
  strata_cols = "DPC",
  nlin = TRUE)
```

```{r}
head(multi_fit2$Parameters)
```

```{r}
multi_fit_K <- fit_multi(
  time_col = "time",
  intensity_col = "random_y",
  data = multi_epidemic,
  strata_cols = "DPC",
  nlin = T,
  estimate_K = T
)
head(multi_fit_K$Parameters)

```

```{r}
multi_fit$Data %>%
  ggplot(aes(time, predicted, color = DPC)) +
  geom_point(aes(time, y), color = "gray") +
  geom_line(size = 1) +
  facet_grid(DPC ~ model, scales = "free_y") +
  theme_minimal()+
  coord_cartesian(ylim = c(0, 1))
```

```{r}
multi_fit$Data %>%
  filter(model == "Gompertz") %>%
  ggplot(aes(time, y=predicted, color = DPC)) +
  geom_point(aes(time, y), size = 2, alpha=.3) +
  geom_line(size = 1.2) +
  theme_minimal() +
  labs(x = "Time", y = "Disease Intensity")
```

Apparent infection rate
The multi_fit$Parameters element is where all stats and parameters as stored. Let’s plot the estimates of the apparent infection rate.

```{r}
multi_fit$Parameters %>%
  filter(model == "Gompertz") %>%
  ggplot(aes(DPC, r)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = r_ci_lwr, ymax = r_ci_upr),
    width = 0,
    size = 1
  ) +
  labs(
    x = "Time",
    y = "Apparent infection rate"
  ) +
  theme_minimal()
```

